---
title: "QuantyFey - Report"
author: "Default User"
output:
  pdf_document: 
    latex_engine: xelatex
  html_document:
    df_print: paged
header-includes:
  - \usepackage{multicol}
  - \setlength{\columnsep}{1cm} % Adjust column spacing
  - \usepackage{caption}
  - \usepackage{booktabs}
  - \usepackage{tabu}
  - \usepackage{longtable} % Required for full-width tables
  - \usepackage{array} % For column alignment customization
  - \setlength{\tabcolsep}{4pt} % Adjust column padding
  - \renewcommand{\arraystretch}{1.2} % Adjust row height
  - \usepackage{pifont}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

knitr::opts_chunk$set(echo = FALSE)
options(knitr.table.format = "latex")

knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

knitr::opts_chunk$set(echo = TRUE)
hook_output <- knitr::knit_hooks$get("output")


theme_set(theme_bw(base_family = "serif"))



```

## Quantitation Summary for `r input$Compound`

This report shows a summary of the quantification of `r input$Compound`. This includes all plots from the overviewiew tap. A summary of the method used for quantification and their individual plots, as well as a summary of all used calibration functions and the used settings.

```{r, echo = F}
table_n = 1
fig_n = 1
```

# Overview of the data
## Retention Time Analysis

The following plot shows the retention time (RT) of the compound for the different sample types. 

```{r, echo = F, fig.height = 3, results = "asis"}
suppressWarnings({
  
  caption_text <- paste0("Figure ", fig_n, ": ", "overviewiew of Retention Time per Sample Type")
  
  fig_n <- fig_n + 1
  
  p <- get_plotly_RT(input, rv)
  p$mapping[["text"]] <- NULL
  #p <- p + labs(x = "Retention Time", y = "Sample Type")+theme(legend.position = "none")
  
  plot(p)

  cat(caption_text)
  cat("\\newline \n")
})

```

\newpage

## Qualifier and Quantifier Analysis
The following plot shows the ratio of the Quantifier and Qualifier transitions (Q/Q).

```{r, echo = F, fig.height = 3, results = "asis"}
suppressWarnings({
  
   caption_text <- paste0("Figure ", fig_n, ": ", "overviewiew of ion ratio per Sample Type")
  
  fig_n <- fig_n + 1
  
  p_quan_qual <- get_plotly_qual_quant(input, rv)
  p_quan_qual$mapping[["text"]] <- NULL
  
  p <- p_quan_qual + labs(y = "Ion Ratio", x = "Sample Type")+ theme(legend.position = "none")
  p <- p 
  
  })

  plot(p)
  
  cat(caption_text)
  cat("\\newline \n")
  
  
```

## Blank Analysis
The following plot shows the blank to sample ratio. 


```{r, echo = F, fig.height = 3, results = "asis"}
suppressWarnings({try({
  
  
   caption_text <- paste0("Figure ", fig_n, ": ", "Overview of Sample to Blank")
  
  fig_n <- fig_n + 1
  
  p <- get_plotly_blank(input, rv)
  p <- p + labs(y = "Blank Ratio", x = "Sample Type")+theme(legend.position = "none") 
  
})

  plot(p)
 
  cat(caption_text)
  cat("\\newline \n")

})
```
\newpage

## Method for Quantification

For this compound the method `r input$quantitation_method` was used. 


```{r, echo = F, warning = F, results = "asis", error = F}




suppressWarnings({
  

  
if(input$quantitation_method == "Custom Bracketing"){
  checkbox_data <- rv$selection_table_bracketing
  
  latex_table <- function(data) {
  # Create LaTeX table header
  header <- paste0(
    "\\begin{tabular}{|c|", paste(rep("c", ncol(data) - 1), collapse = "|"), "|}\n",
    "  \\hline\n  ", 
    paste(colnames(data), collapse = " & "), " \\\\\n  \\hline"
  )
  
  # Create rows
  rows <- apply(data, 1, function(row) {
    item <- row[1] # First column is the row label (e.g., "Item")
    options <- sapply(row[-1], function(x) {
      if (x == TRUE) {
        "$\\boxtimes$"  # Checked box
      } else {
        "$\\Box$"       # Unchecked box
      }
    })
    paste0(item, " & ", paste(options, collapse = " & "), " \\\\")
  })
  
  # Combine everything into a LaTeX table
  table <- paste0(
    header, "\n",
    paste(rows, collapse = "\n"), "\n",
    "  \\hline\n\\end{tabular}"
  )
  return(table)
}

  cat("The following table shows the Calibration Functions used for the specific Blocks within the Measured Sequence.")
  cat(paste0("\\newline \n\n Table ", table_n, ": Bracketing Table used for the analysis."))
  
  table_n <- table_n + 1
  
# Print LaTeX table
cat(latex_table(checkbox_data))
  
}else if(input$quantitation_method == "IS Correction"){
  cat("The red dots mark the Peak Areas of the Internal Standard compounds and the Bars represent the Peak Areas of the Compound over the whole Measurement Sequence. The lower plot represents the ratios from the respective Sample after each Peak Area from the Compound was divided by the Peak Area from the selected Internal Standard.\\newline \n\n")
  
  # Combine plots using patchwork

tryCatch({
  combined_plot <- (rv$IS_plot[[1]] + 
      theme(legend.position = "none")) / 
      (rv$IS_plot[[2]] + 
      theme(legend.position = "bottom"))  # Stack plots vertically

  # Display the combined plot
  print(combined_plot)

  caption_text <- paste0("Figure ", fig_n, ": ", "Overview of internal standard correction")
  cat(caption_text)
  cat("\\newline \n")
}, error = function(e) NA, warning = function(w) NA)


  cat("\\newline \n")
  cat("The following table shows the correction factors used for the specific Sample Types. This is useful if different dilutions were used for the analysis.")
  cat(paste0("\\newline \n Table ", table_n, ": Information about the correction factors for each sample type"))
 
  print(kable(as.data.frame(rv$IS_table), format = "latex", row.names = F, booktabs = T))



}else if(input$quantitation_method == "Drift Correction"){
  cat("The following plots showcase the Peak Areas for the selected compound from the measured data. The first plot show the Peak Areas as they were processed without any manipulation with red indicated bars showcasing the Samples selected for Drift Correction and the red line indicating the model used for Drift Correction. The second plot shows the manipulated data after Drift Correction.\\newline \n")
  
  p1 <- rv$plot_dc[[1]] + theme(legend.position = "none") 
  p2 <- rv$plot_dc[[2]] + theme(legend.position = "bottom")

  legend <- get_legend(p2)

  p2_clean <- p2 + theme(legend.position = "none")

  plots <-  plot_grid(p1, p2_clean, nrow = 2, align = "hv")
  plot_grid(plots, legend,
  rel_heights = c(1, 0.1),
  nrow = 2
) %>% plot()

  caption_text <- paste0("Figure ", fig_n, ": ", "Overview of Drift Correction")
  cat(caption_text)
  cat("\\newline \n")
  
}else if(input$quantitation_method == "Default Bracketing"){
  
  df <- rv$data[,c("Sample.Name","Sample.Type", input$Compound)]
  colnames(df)[colnames(df) == input$Compound] <- "Peak Area"
  df$inj <- 1:nrow(df)
  
  p <- ggplot(df, aes(x = inj, y = `Peak Area`, fill = Sample.Type))+geom_bar(stat = "identity", col = "black") + 
  scale_fill_manual(values = c("Model" = "red3", "Sample" = "lightblue3", "Cal" = "navy", "Blank" = "grey70", "QC" = "purple3"))
  
  plot(p)

  caption_text <- paste0("Figure ", fig_n, ": ", "Sequence Overview")
  fig_n <- fig_n + 1
  cat(caption_text)
  cat("\\newline \n")
}

})

```




```{r, echo = F, warnings = F, error= F}
    quantitate_data <- quantitate(input, rv)
    quant_data <- quantitate_data[[1]]
    mods <- quantitate_data[[3]]
               
    
                               
    plots_quant <- list()
    for(block_name in names(rv$selection_cals_table)){
    
    mod <- mods[[block_name]]

    cals_data_block <- rv$selection_cals_table[[block_name]]
     quant_data_block <- quant_data[quant_data$Classification == block_name,]
     quant_data_not_block <- quant_data[!quant_data$Classification == block_name,]

     p <- generate_quantitate_ggplot(data = cals_data_block, regression_model = mod, regression_method = input$regression_model)
    
    p <- p + theme(plot.title = element_blank())

      
      plots_quant[[block_name]] <- p
      
    }
```



```{r, echo = F, warnings = F, error = F}
   
suppressWarnings({
  

    mods <- quantitate_data[[3]]
    
    results_acc <- list()
    
    for(block in names(rv$selection_cals_table)){
      
      
      
      df <- quantitate_data[[1]] 
    
    mod_temp <- mods[[block]]
    used_cals <- unique(rv$selection_cals_table[[block]]$Classification)

    
    #print(used_cals)
    pd <- pd_temp(rv)

    df <- df[(df$Classification %in% used_cals) | ((df$Classification == block) & (pd == "Cal")),]

    df$Concentration <- sapply(df$Sample.Name, FUN = function(x, label, concentration){
      for(i in 1:length(label)){
        if(grepl(label[i], x)){
          return(concentration[i])
        }
      }
    }, label = rv$setup_cal$Cal.Name,
    concentration = rv$setup_cal$Concentration)

    

    df$pred <- predict_concentrations(data = df, model = mod_temp ,method = input$regression_model)

    df$pred <- signif(df$pred, digits = 3)



    df$PeakArea <- round(df$PeakArea, 1)

    df$Accuracy <- signif(df$pred / df$Concentration * 100, 3)

    df <- dplyr::select(df, Sample.Name, Classification, PeakArea, Concentration, pred, Accuracy)

    
    results_acc[[block]] <- df
    }
    
})

```


\newpage

# Quantitation Summary
## Summary of the Calibration Functions and the Accuracy of the models

The following plots show for all separate Blocks the Calibration model used for the specific blocks. 


```{r, results = "asis", warnings = F, echo = F, fig.height = 4}


auto_set_format <- function() { 
   if (knitr::is_latex_output()) { 
     options(knitr.table.format = "latex") 
   } else { 
     options(knitr.table.format = "html") 
   } 
 }

 
escape_md <- function(text) {
  # Escape special Markdown characters, including '+'
  gsub("([\\*\\_\\[\\]\\(\\)\\{\\}<>`#\\+\\-!|])", "\\\\\\1", text, perl = TRUE)
}


comment_escaped <- escape_md(input$Comment)

auto_set_format()


# Suppress warnings to avoid unnecessary output
 suppressWarnings({
   # Loop through each block in the calibration plots
   for (i in names(plots_quant)) {
  
     # Skip blocks that contain "Cal" in their name
     if (grepl("Cal", i)) {
       next
     }
  
     # Convert the results for the current block into a data frame
     dt <- as.data.frame(results_acc[[i]])
  
     # Generate a caption for the calibration plot
     caption_text <- paste0("Figure ", fig_n, ": ", "Calibration Plot for Block: ", i, sep = "")
     fig_n <- fig_n + 1
  
     # Add the caption to the plot and adjust its theme
     p <- plots_quant[[i]]# +

      #cat("\\newline \n")
      



       #theme(plot.caption = element_textbox_simple(padding = margin(0, 10, 5, 0)))
  
     # Print the plot
     print(p)

     cat(caption_text)
    cat("\\newline \n")
  
     # Generate specific settings for the quantitation method
     quantitation_method_specifific <- sapply(input$quantitation_method, FUN = function(x) {
       if (x == "Drift Correction") {
         return(paste0(
           "Model for Drift Correction: ", ifelse(input$model_drift == "lm", "linear", "LOESS"), "\\newline \n",
           "File for Drift Correction: ", input$files_for_correction, 
           ifelse(input$model_drift != "lm", paste0("\\newline \n", "Span Width: ", input$span_width), "")
         ))
       } else if (x == "IS Correction") {
         compound_IS <- input$Compound_IS %>% gsub("_", "-", .)
         return(paste0("IS Compound for Correction:\\newline \n", compound_IS))
       } else {
         return(c(""))
       }
     })
  
     # Start a two-column layout for the summary
     cat("\\begin{multicols}{2}")
  
     # Add a table for concentration levels
     cat(paste0("Table ", table_n, ": ", "Concentration levels ", "\\newline"))
     print(kable(rv$setup_cal, booktabs = T))
     table_n <- table_n + 1
  
     # Add a column break for the second column
     cat("\\columnbreak \\newline \n")
  
     # Add a summary of settings
     cat("\\textbf{Summary of Settings:}\\newline")
     cat(paste0(
       "\\newline \n",
       "Model used: ", input$regression_model, "\\newline \n",
       "Quantitation Method used: ", input$quantitation_method, "\\newline \n",
       "Lower Limit of Quantification: ", rv$LLOQs[[i]], "\\newline \n",
       "Weighting Method: ", input$weight_method, "\\newline \n",
       quantitation_method_specifific, "\\newline \n",
       "Comment added to this Compound:\\newline \n  ", comment_escaped, "\\newline \n", "\\newline \n"
     ))
  
     # End the two-column layout
     cat("\\end{multicols}")
  
     # Add a table for accuracy results
     cat(paste0("\n\n\\pagebreak\n", "Table ", table_n, ": Accuracy Table of Block: ", i, "\\newline \n"))
     table_n <- table_n + 1
     print(kable(dt, row.names = F, booktabs = T) %>% kable_styling(full_width = T))
  
     # Break the loop if the block is "all"
     if (i == "all") {
       break
     }
  
     # Add a page break for the next block
     cat("\n\n\\pagebreak\n")
   }
 })
```



